---
title: "Bureau of Economic Analysis | Infrastructure Investment"
author: "Alexandria Hughes"
date: "8/16/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

## Libraries

```{r}
library(tidyverse)
library(scales)
library(ggridges) #for beautiful ridgeline plots
```

## Read in the data

These data come from a working paper from [Jennifer Bennett et al. at the Bureau of Economic Analysis](https://www.bea.gov/system/files/papers/BEA-WP2020-12.pdf) and contain investment in basic, social, and digital infrastructure in the United States National Economic Accounts from 1947 - 2017.

```{r}
investment <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-10/investment.csv')
chain_investment <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-10/chain_investment.csv')
ipd <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-10/ipd.csv')
```

#### Initial exploration

Three datasets are provided:

  + investment: values in millions
  + chain_investment: same as Investment but chained to 2021 dollars
  + ipd
  
```{r}
str(chain_investment)
```

Broken up into categories and meta-categories. It looks like meta-cats will be easier to work with since there are fewer of them and the naming is still informative.
```{r}
#summary(factor(investment$category))
summary(factor(investment$meta_cat))
```



## Visualization

I'm not an economist, but I think chain_investment will be easier to think about and make visual comparisons with since it's already normalized. We'll move forward with the chain_investment data.

#### Create some year bins

chain_investment has entries for each year, 1947-2017. I think it will be useful to break this range up into larger segments of time. A good way to do this generally is to break up a range with `cut()`. However, I wanted nice decade intervals (rather than exact 1947-1957, etc) so I created a new variable, `year_range`, that assigns each year to a ~nice~ decade (the ends contain a little more/little less than a true decade).

```{r}
chain_investment <- chain_investment%>% 
  mutate(year_range = case_when(
            between(year, 1946,1959) ~ "1947-1959",
            between(year, 1960,1969) ~ "1960-1969",
            between(year, 1969,1979) ~ "1970-1979",
            between(year, 1979,1989) ~ "1980-1989",
            between(year,1989,1999) ~ "1990-1999",
            between(year,2000,2009) ~ "2000-2009",
            between(year,2010,2018) ~ "2010-2017"))
```


### Social investment

The social meta-category contains categories Education, Federal, Health, Private, Public Safety, and S&L. Here is a plot of the social meta-category, with chained gross investment on the x axis, and faceted (broken up into individual plots) for each decade. I like this because you can see the relationship between the categories over the decades! Notice how Private and Health (cyan and green) advance relative to Public Safety and Federal, which are relatively stable over time (but do note that each decade has its own x-axis!).

```{r}
ggplot(chain_investment %>% filter(meta_cat == "Social"), aes(x=gross_inv_chain, y=category, fill=category))+
  theme_ridges(font_size = 10)+
  facet_wrap(~year_range, scales = "free_x", nrow = 2)+
  geom_density_ridges(
    jittered_points = TRUE,
    position = position_points_jitter(width = 0.5, height = 0),
    point_shape = "|", point_size = 1.5, point_alpha = 1, alpha = 0.5)+
  labs(x="Gross Investment (Millions of chained 2021 dollars)", 
       y="", 
       title = "Federal investment in social infrastructure 1947-2017", 
       caption = "Alex Hughes | Data courtesy of Bureau of Economic Analysis (BEA)")+
  xlim(0,100000)+
  scale_x_continuous(expand = c(0.125,0.125), n.breaks = 3, labels = scales::comma)+
    theme(legend.position = "none",
        legend.text = element_text(size=10),
        plot.title = element_text(size=12),
        plot.subtitle = element_text(size=12),
        plot.caption = element_text(size=8))
```

If we drop the facets, this becomes even more clear - at the cost of seeing each decade as clearly.

```{r}
ggplot(chain_investment %>% filter(meta_cat == "Social"), aes(x=gross_inv_chain, y=year_range, fill=category))+
  theme_ridges(grid = FALSE, font_size = 10)+
  geom_density_ridges(
    alpha = 0.5)+
  theme(legend.position = "top")+
  labs(x="Gross Investment (Millions of chained 2021 dollars)", 
       y="Year Range", 
       fill="",
       title = "Federal investment in social program categories 1947-2017", 
       subtitle = "Health & Private spending split away from the pack", 
       caption = "Alex Hughes | Data courtesy of Bureau of Economic Analysis (BEA)")

```

### Electric power

Now let's try electric power! I think this will be interesting because one of the contained categories is wind and solar power investment, which I would predict (as a total power layperson who is at least aware that these are renewable energy sources) has seen increased investment over time. 

```{r}
ggplot(chain_investment %>% filter(meta_cat == "Electric power"), aes(x=gross_inv_chain, y=year_range, fill=category))+
  theme_ridges(grid = FALSE, font_size = 10)+
  stat_density_ridges(
    quantile_lines = TRUE, 
    quantiles = 2,
    alpha = 0.5)+
  theme(legend.position = "top",
        legend.text = element_text(size=10),
        plot.title = element_text(size=14),
        plot.subtitle = element_text(size=12),
        plot.caption = element_text(size=10))+
  guides(fill = guide_legend(nrow = 3))+
  labs(x="Gross Investment (Millions of chained 2021 dollars)", 
       y="Year Range", 
       fill="",
       title = "Federal investment in electric power 1947-2017", 
       subtitle = "Recent decades see increasing investment in wind and solar power structures", 
       caption = "Alex Hughes | Data courtesy of Bureau of Economic Analysis (BEA)")
```
```{r}
ggplot(chain_investment %>% filter(meta_cat == "Electric power"), aes(x=gross_inv_chain, y=category, fill=category))+
  theme_ridges(font_size = 10)+
  facet_wrap(~year_range, scales = "free_x", nrow = 2)+
  geom_density_ridges(
    jittered_points = TRUE,
    position = position_points_jitter(width = 0.5, height = 0),
    point_shape = "|", point_size = 1.5, point_alpha = 1, alpha = 0.5)+
  labs(x="Gross Investment (Millions of chained 2021 dollars)", 
       y="", 
       title = "Federal investment in electric power 1947-2017", 
       caption = "Alex Hughes | Data courtesy of Bureau of Economic Analysis (BEA)")+
  xlim(0,100000)+
  scale_x_continuous(expand = c(0.125,0.125), n.breaks = 3, labels = scales::comma)+
    theme(legend.position = "none",
        legend.text = element_text(size=10),
        plot.title = element_text(size=12),
        plot.subtitle = element_text(size=12),
        plot.caption = element_text(size=8))
```



### Education

Finally, we can take similar approaches to visualize infrastructure investment in education.

```{r}
ggplot(chain_investment %>% filter(meta_cat == "Education"), aes(x=gross_inv_chain, y=category, fill=category))+
  theme_ridges(font_size = 10)+
  geom_density_ridges(
    alpha = 0.5)+
  scale_fill_discrete(labels = c("Federal", "Private", "S&L Higher Ed", "S&L K-12", "S&L Libraries", "S&L Other Ed Structures"))+
  facet_wrap(~year_range, scales = "free_x", nrow = 2)+
  theme(legend.position = "none",
        legend.text = element_text(size=10),
        plot.title = element_text(size=12),
        plot.subtitle = element_text(size=12),
        plot.caption = element_text(size=10),
        text = element_text(size=12))+
  scale_x_continuous(n.breaks = 3, expand = c(0.1,0.1), labels = scales::comma)+
  labs(x="Gross Investment (Millions of chained 2021 dollars)", 
       y="",
       title = "Federal investment in education 1947-2017",
       caption = "Alex Hughes | Data courtesy of Bureau of Economic Analysis (BEA)")
  
```


```{r}
ggplot(chain_investment %>% filter(meta_cat == "Education"), aes(x=gross_inv_chain, y=year_range, fill=category))+
  theme_ridges(grid = FALSE, font_size = 10)+
  stat_density_ridges(
    quantile_lines = TRUE, 
    quantiles = 2,
    alpha = 0.5)+
  theme(legend.position = "top",
        legend.text = element_text(size=10),
        plot.title = element_text(size=14),
        plot.subtitle = element_text(size=12),
        plot.caption = element_text(size=10),
        text = element_text(size=12))+
  guides(fill = guide_legend(nrow = 3))+
  labs(x="Gross Investment (Millions of chained 2021 dollars)", 
       y="Year Range", 
       fill="",
       title = "Federal investment in education 1947-2017", 
       subtitle = "Library & structural investment steady, while K-12 & higher ed soar",
       caption = "Alex Hughes | Data courtesy of Bureau of Economic Analysis (BEA)")
```






